---
title: "Untitled"
output: word_document
---


```{r}
library(tidyverse)
library(lubridate)
library(covid19ita)
library(scales)
```

```{r}
load(here::here("data", "db_mortality_rev.rda"))
load(here::here("data", "veneto_res_rev.rda"))
data("dpc_covid19_ita_regioni")
```

```{r include = FALSE}

#if in a day there are no deaths in a municipality, the line
#for that day is missing. I add it 

create_date_utili <- function(data1, data2){
    
    data.frame(date = seq(as_date(data1),
                     as_date(data2), "days"))
}

date_utili <- map2_df(.x = c("2015-01-01", "2016-01-01", "2017-01-01",
                          "2018-01-01", "2019-01-01", "2020-01-01"),
                   .y = c("2015-06-30", "2016-06-30", "2017-06-30",
                          "2018-06-30", "2019-06-30", "2020-06-30"),
                   ~create_date_utili(.x, .y))

    
    
    
db_veneto_wide <- db_mortality_rev %>% 
    filter(nome_regione == "Veneto") %>% 
    mutate(comune = as.character(nome_comune)) %>% 
    dplyr::select(-reg, -prov, -nome_regione, -cod_provcom,
                  -nome_comune) %>% 
    unite(ggmm, gg, mm, sep = "-") %>% 
    filter(is.finite(t_20)) 
    
    
db_veneto_long <- db_veneto_wide %>% 
    pivot_longer(cols = m_15:t_20, 
                 names_to = "annosesso",
                 values_to = "morti") %>% 
    separate(annosesso, into = c("sesso", "anno"), sep = "_") %>% 
    unite(date, ggmm, anno, sep = "-") %>% 
    mutate(date = dmy(date)) %>% 
    arrange(date) %>% 
    group_by(comune, sesso, cl_eta) %>% 
    right_join(date_utili) %>% 
    ungroup()


    
```

```{r include = FALSE}

seq_gg <- data.frame(seq_gg = rep(1:181, 6))

db_plot1 <- db_veneto_long %>% 
    filter(sesso == "t") %>% 
    group_by(date) %>% 
    mutate(morti_tot = sum(morti)) %>% 
    ungroup() %>% 
    distinct(date, .keep_all = TRUE) %>% 
    filter(date != "2020-02-29" & date != "2016-02-29") %>% 
    separate(date, into = c("aa", "mm", "gg"), sep = "-", 
             remove = FALSE) %>% 
    bind_cols(seq_gg)


```


```{r include = FALSE}

res_db <- left_join(db_veneto_wide, veneto_res_rev)


res_long <- res_db %>% 
    dplyr::select(-(m_15:t_20), -x2) %>% 
    distinct(comune, .keep_all = TRUE) %>% 
    pivot_longer(cols = c(maschi15:totale20),
                 names_to = "annosesso",
                 values_to = "residenti") 


res_long$as <- as.character(strsplit(res_long$annosesso, "1"))
res_long$as <- NA

for(i in 1:nrow(res_long)) {
    
    if(str_detect(res_long$annosesso[i], pattern = "1")) {
        
        res_long$as[i] <- as.character(strsplit(res_long$annosesso[i], "1"))
    } else {
        
        res_long$as[i] <- as.character(strsplit(res_long$annosesso[i], "2"))
    }
}


res_long1 <- res_long %>%
    separate(as, into = c("sesso", "anno"), sep = ",",
             remove = FALSE) %>% 
    mutate(sesso = ifelse(sesso == 'c("femmine"', "femmine",
                          ifelse(sesso == 'c("maschi"', "maschi",
                                 ifelse(sesso == 'c("totale"', "totale",
                                        sesso))),
           anno = ifelse(anno == ' "5")', 2015,
                     ifelse(anno == ' "6")', 2016,
                             ifelse(anno == ' "7")', 2017,
                                    ifelse(anno == ' "8")', 2018,
                                        ifelse(anno == ' "9")', 2019,
                                               ifelse(anno == ' "0")', 2020,
                                                      NA))))))) %>% 
    group_by(annosesso, provincia) %>%
    mutate(res_tot = sum(residenti, na.rm = TRUE)) %>%
    ungroup() %>%
    distinct(annosesso, provincia, .keep_all = TRUE) %>%
    dplyr::select(-as, -annosesso, -ggmm, -cl_eta)



```


plot 1

```{r include = FALSE}


tasso_mort1 <- res_long1 %>% 
    filter(sesso == "totale") %>% 
    mutate(aa = as.character(anno)) %>% 
    group_by(aa) %>% 
    mutate(res_veneto = sum(res_tot)) %>% 
    ungroup() %>% 
    distinct(aa, .keep_all = TRUE) %>% 
    right_join(db_plot1, by = "aa") %>% 
    right_join(date_utili) %>% 
    group_by(aa) %>% 
    mutate(tasso = (morti_tot/res_veneto)*100000) %>% 
    filter(date != "2020-02-29" & date != "2016-02-29") %>% 
    mutate(anno_falso = 1888) %>% 
    separate(date, into = c("yearr", "mese", "giorno"), sep = "-",
             remove = FALSE) %>% 
    unite(date_false, anno_falso, mese, giorno, sep = "-",
          remove = FALSE) %>% 
    mutate(date_false = as_date(date_false)) %>% 
    mutate(year = ifelse(yearr == "2020", "2020", "average 2015-2019")) %>%
    unite(data_falsissima, year, mese, giorno, remove = FALSE) %>% 
    group_by(data_falsissima) %>% 
    mutate(average_rate = mean(tasso),
           year = as.factor(year)) %>% 
    ungroup() %>% 
    distinct(data_falsissima, .keep_all = TRUE)




```

```{r echo = FALSE}
plot1 <- ggplot(tasso_mort1) +
    geom_smooth(aes(date_false, average_rate, colour = year), 
                span = 0.5, se = T)+
    xlab("Date") +
    ylab("Daily mortality rate x 100000 inhabitants")  +
  theme_bw() +
  theme_minimal() +
  geom_vline(aes(xintercept = as.numeric(date_false)[67])) +
  scale_x_date(breaks = "month", labels = label_date_short())
  

# ggsave("plot1.tiff", units="in",  dpi=500, compression = 'lzw',
#        path = here::here("img"), width = 9, height = 5)
```



Mortality rate in males, females and overall in march and april 2019 
and 2020 stratified by province resulting from the aggregation of selected municipalities. Mortality change rate in march and april from 2019 to 2020 in males, females and overall.
```{r include = FALSE}

db_var_sex_tax <- res_db %>% 
    separate(ggmm, into = c("gg", "mm"), sep = "-", remove = FALSE) %>% 
    filter(mm == "3" | mm == "4") %>% 
    group_by(comune) %>% 
    mutate(m19 = sum(m_19),
           m20 = sum(m_20),
           f19 = sum(f_19),
           f20 = sum(f_20),
           t19 = sum(t_19),
           t20 = sum(t_20)) %>% 
    ungroup() %>% 
    distinct(comune, .keep_all = TRUE) %>% 
    group_by(provincia) %>% 
    mutate(m_sum19 = sum(m19),
           m_sum20 = sum(m20),
           f_sum19 = sum(f19),
           f_sum20 = sum(f20),
           tot_sum19 = sum(t19),
           tot_sum20 = sum(t20),
           res_m19 = sum(maschi19, na.rm = TRUE),
           res_m20 = sum(maschi20, na.rm = TRUE),
           res_f19 = sum(femmine19, na.rm = TRUE),
           res_f20 = sum(femmine20, na.rm = TRUE),
           res_tot19 = sum(totale19, na.rm = TRUE),
           res_tot20 = sum(totale20, na.rm = TRUE)) %>% 
    distinct(provincia, .keep_all = TRUE) %>% 
    mutate(
      var_m = (((m_sum20/res_m20) - (m_sum19/res_m19))/
                   (m_sum19/res_m19))*100,
      var_f = (((f_sum20/res_f20) - (f_sum19/res_f19))/
                   (f_sum19/res_f19))*100,
      var_tot = (((tot_sum20/res_tot20) - (tot_sum19/res_tot19))/
                   (tot_sum19/res_tot19))*100,
      tasso_m20 = (m_sum20/res_m20)*1000,
      tasso_m19 = (m_sum19/res_m19)*1000,
      tasso_f19 = (f_sum19/res_f19)*1000,
      tasso_f20 = (f_sum20/res_f20)*1000,
      tasso_t19 = (tot_sum19/res_tot19)*1000,
      tasso_t20 = (tot_sum20/res_tot20)*1000)
    

```

```{r}
data.frame(provincia = db_var_sex_tax$provincia,
           tasso_m19 = db_var_sex_tax$tasso_m19,
           tasso_m20 = db_var_sex_tax$tasso_m20,
           tasso_f19 = db_var_sex_tax$tasso_f19,
           tasso_f20 = db_var_sex_tax$tasso_f20,
           tasso_t19 = db_var_sex_tax$tasso_t19,
           tasso_t20 = db_var_sex_tax$tasso_t20,
           var_m = db_var_sex_tax$var_m,
           var_f = db_var_sex_tax$var_f,
           var_tot = db_var_sex_tax$var_tot) %>% 
    knitr::kable(digits = 2L)
```




number of deaths in the Veneto Region resulting from the aggregation of selected municipalities from January 1st 2020 to July 30th 2020 stratified by classes of age


```{r include = FALSE}


db_andamento_priv <- db_veneto_long %>% 
    filter(sesso == "t" & date >= "2020-01-01") %>% 
    mutate(cl_eta = ifelse(cl_eta %in% c(0:13), "0-65",
               ifelse(cl_eta %in% c(14, 15), "66-75",
                ifelse(cl_eta %in% c( 16:21), "76 +", cl_eta)))) %>% 
    group_by(provincia, date) %>% 
    mutate(morti_tot = sum(morti)) %>% 
    ungroup() %>% 
    group_by(provincia, date, cl_eta) %>% 
    mutate(mort_cl_eta = sum(morti)) %>% 
    ungroup() %>% 
    distinct(provincia, date, cl_eta, .keep_all = TRUE) %>% 
    mutate(`Age class` = cl_eta)


```

```{r fig.height=4, fig.width=8, echo = FALSE}
plot2 <- db_andamento_priv %>% 
    group_by(date, cl_eta) %>% 
    mutate(m_veneto = sum(mort_cl_eta)) %>% 
    distinct(date, cl_eta, .keep_all = TRUE) %>% 
ggplot() +
    geom_line(aes(date, m_veneto, colour = `Age class`), size = 1) +
    ylab("Number of deaths") +
    xlab("Date") +
     theme_bw() +
    theme_minimal() +
  scale_x_date(breaks = "month", labels = label_date_short())


# ggsave("plot2.tiff", units="in",  dpi=500, compression = 'lzw',
#        path = here::here("img"), width = 9, height = 5)
```



```{r}


#focus on march-april, months of lockdown
covid_veneto3_4 <- dpc_covid19_ita_regioni %>% 
    filter(denominazione_regione == "Veneto") %>% 
    separate(data, into = c("day", "time"), remove = FALSE, sep = " ") %>% 
    filter(day > "2020-02-28" & day < "2020-05-01") #deaths 1459(tot deceduti 
#2020-04-30) - 2 (deceduti before 2020-03-01) = 1457

(1457/4907704)*100000 #covid mortality in march and april
#29.7

a3_4 <- db_veneto_long %>% 
    filter(sesso == "t") %>% 
    filter(date > "2020-02-29" & date < "2020-05-01") %>% 
    separate(date, into = c("year", "month", "day"), remove = FALSE) %>% 
    group_by(year) %>% 
    mutate(deaths = sum(morti)) %>% 
    ungroup() %>% 
    distinct(year, .keep_all = TRUE)


(9993/4560246)*100000 #all-causes mortality in march and april
#219.1


b3_4 <- db_veneto_long %>% 
    filter(sesso == "t") %>% 
    separate(date, into = c("year", "month", "day"), remove = FALSE) %>% 
    filter(month == "03" | month == "04") %>% 
    group_by(year) %>% 
    mutate(deaths = sum(morti)) %>% 
    ungroup() %>% 
    distinct(year, .keep_all = TRUE)

((8031+7663+7820+7990+7637) / (4528822+4516950+4519599+4520610+4522371))*100000

#173.1


219.1-173.1 # = 46
```







